### 2015-06-14 lp.tex
LUB cleanup

The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by adding the extra constraint $J=\Phi r$ in \eqref{mdplpshort} with $\Phi \in \R^{n\times k}$ and the new variables $r\in \R^k$. 
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\\
\text{s.t.}\mb &\Phi r\geq T \Phi r.
\end{split}
\end{align}
Unless specified otherwise we use $\tr$ to denote any solution\todoc{instead of saying the solution} to the ALP and $\tj=\Phi \tr$ to denote the corresponding approximate value function. 
The extra constraint added that $J$ lies in the (column) span of $\Phi$ could lead to an infeasible LP. An easy way to prevent this is if the vector $\one$ whose components are all equal to one is in the span of $\Phi$.
Indeed, by \cref{shift}, for any $t\in \R$, $T (t\one ) =  \alpha t \one+T \zero \le \alpha t \one + \max_{s,a} g_a(s)$, hence $t\one \ge \alpha t \one+T\zero $ happens as soon as
$t\ge \frac{1}{1-\alpha}\max_{s,a} g_a(s)$. 
This leads to the following assumption:
\begin{assumption}\label{one}
The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$. 
In other words, the constant function is part of the basis.
\end{assumption}

>>>>

The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by adding the extra constraint $J=\Phi r$ in \eqref{mdplpshort} with $\Phi \in \R^{n\times k}$ and introducing 
the new variables $r\in \R^k$. 
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\\
\text{s.t.}\mb &\Phi r\geq T \Phi r.
\end{split}
\end{align}
Unless specified otherwise we use $\tr$ to denote any solution 
to the ALP and $\tj=\Phi \tr$ to denote the corresponding approximate value function. 
The extra constraint added that $J$ lies in the (column) span of $\Phi$ could lead to an infeasible LP. An easy way to prevent this is if the vector $\one$ whose components are all equal to one is in the span of $\Phi$.
Indeed, by \cref{shift}, for any $t\in \R$, $T (t\one ) =  \alpha t \one+T \zero \le \alpha t \one + \max_{s,a} g_a(s)$, hence $t\one \ge \alpha t \one+T\zero $ happens as soon as
$t\ge \frac{1}{1-\alpha}\max_{s,a} g_a(s)$. 
This leads to the following assumption:
\begin{assumption}\label{one}
The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$. 
In other words, the constant function is part of the basis.
\end{assumption}
Under this assumption \eqref{alp} is thus feasible and since \eqref{alp} is obtained by adding an extra constraint to the LP \eqref{mdplpshort}, the solutions to  \eqref{alp} are bounded and $\tj$ is uniquely defined.

### 2015-06-14 lubp.tex
LUB cleanup

@@

Given $J\in \R^n$, $c\in \R^n$ positive-valued vector define $r_{c,J}$ to be the solution to 

@@

 \underset{r\in \N}{\min} &\,\, c^\top \Phi r\,,\\
 \text{s.t.}& \mb \Phi r\geq  TJ\,,

@@

where it is assumed that a (bounded) solution exists.

@@

@@

\item Given $J\in \R^n$, define $\F_J\eqdef\{\,\Phi r\,|\,\Phi r\geq TJ, r\in \N\,\}$. 

@@

Further, since $(\Gamma J)(i) = \min\{ V(i) \,|\, V\in \F_J \}$, it also follows that $\Gamma J \le V$ for any $V\in \F_J$.

@@

\begin{definition}\label{bestproj}
The LUB projection of $J^*$ is denoted by $\bj=\Gamma J^*$. \todoc{It is a bit strange to call this a definition. A numbered equation could be enough.}

@@

As mentioned earlier, the error analysis depends on two $\max$-norm contraction operators the first of which is $\Gamma$. The important result of this section is Theorem~\ref{fxpres} and it relates the fixed point $\tv$ of $\Gamma$ to $J^*$.
\todoc{Poor choice of symbols: $\tv$ is the fixed point of $\Gamma$ and $\hat{V}$ will be the fixed point of $\tilde{\Gamma}$? Hmm.. Not a high priority to change it, but still..}
In what follows, we let $r^*\in \R^k$ be defined as $r^*\eqdef\argmin_{r\in R^k}||J^*-\Phi r||_\infty$.

@@

\begin{lemma}\label{gmonotone}
For $J_1, J_2\in \R^n$ such that $J_1\geq J_2$, we have $\Gamma J_1\geq \Gamma J_2$.
\end{lemma}
\begin{proof}
Choose any $i\in \{1,\ldots,n\}$. 
Since $J_1\geq J_2$, we have $TJ_1\geq TJ_2$.
Hence, $\F_{J_1} \subset \F_{J_2}$ and thus $(\Gamma J_1)(i) \ge (\Gamma J_2)(i)$. 
Since $i$ was arbitrary, the result follows.
\end{proof}
\begin{lemma}\label{lpsol}
Let $A\in \R^{u\times v}$, $b,c\in \R^u$ and $b_0=Ax_0$ for 
some $x_0 \in \R^v$, $\N \subset \R^v$. Then
\begin{align}
\min\{c^\top Ax:Ax\geq b+b_0, x\in \N\} =\min\{c^\top A y:Ay \geq b, y+x_0 \in \N \}+c^\top b_0.
\end{align}
\end{lemma}
\begin{proof}
The claim follows by the change of variables $y := x-x_0$.
\end{proof}
\begin{lemma}\label{gshift}
Let $J\in \R^n$ and $t\in \R$ be a constant.
Then, $\Gamma (J+t\one) = \Gamma J + \alpha t \one$ provided that $r\in \N$ holds if and only $r+s e_1\in \N$ holds for any $s\in \R$.
\end{lemma}
\begin{proof}
Consider the $i^{th}$ linear programs associated with $\Gamma J$ and $\Gamma (J+ t\one)$. 
The result follows by using Lemma~\ref{lpsol} with $A=\Phi$, $b=TJ$, $c=e_i$, $b_0=\alpha t\mathbf{1}$ 
and $x_0=\alpha t e_1$, noting that by our assumption on $\N$,
$\{ y\,: \, \Phi y \geq TJ, y + \alpha t e_1\in \N \}
= \{ y\,: \, \Phi y \geq TJ, y \in \N \}$. \todoc{This is where this assumption is needed.}
\end{proof}
\begin{theorem}\label{gmaxcontra}
The operator $\Gamma  \colon \R^n\ra \R^n$ obeys the $\max$-norm contraction property with factor $\alpha$.
\end{theorem}
\begin{proof}
The statement follows immediately from Lemmas~\ref{gmonotone} and~\ref{gshift}, and \cref{maxnorm}.
\end{proof}
\if0
\begin{corollary}
Let $V_0$ be arbitrary and $(V_n)$ be defined recursively by
\begin{align}\label{pvi}
V_{n+1}&=\Gamma V_n, \quad n\geq 0\,.
\end{align}
Then, $V_n$ converges to the the unique fixed point $\tv$ of $\Gamma$.
\end{corollary}
\fi
Define $\tv$ to be the unique fixed point of $\Gamma$.
\begin{lemma}\label{gfp}
The vector $\tv$ obeys $\tv\geq T\tv$.
\end{lemma}
\begin{proof}
Let $1\le i \le n$. We have $\tv(i) = (\Gamma \tv)(i) = (\Phi r_{e_i,\tv})(i) \ge (T\tv)(i)$.
\if0
Consider the $i^{th}$ linear program associated with $\Gamma \tv$. We know that $\Phi r_{e_i}\geq T \tv$, for all $i=1,\ldots, n$. The result follows from noting that $\tv$ is the unique fixed point of $\Gamma $ and that $\tv(i)=(\Phi r_{e_i,V})(i)$.
\fi
\end{proof}
Recall that $\tj$ denotes the solution to the ALP in \eqref{alp}.
\begin{lemma}\label{relation1}
Assume that $\N$ is such that if $r_i$ is
the solution to the ALP in \eqref{alp} for $c=e_i$ then $r_i\in \N$, $1\le i \le n$.
% $\tv$, the unique fixed point of the iterative scheme \eqref{pvi}, and the solution $\tj$ to the ALP in \eqref{alp}, obey 
Then, the relation $\tj\geq\tv\geq J^*$ holds.
\end{lemma}
\begin{proof}
Since $\tv\geq T\tv$ it follows that $\tv\geq J^*$. 
%Let $\Phi r_1, \Phi r_2,\ldots,\Phi r_n$ be solutions to the ALP in \eqref{alp} for $c=e_1, e_2,\ldots,e_n$ respectively. 
Define $V_0\in \R^n$ by $V_0(i)=\underset{j=1,\ldots, n}{\min}(\Phi r_j)(i)$, $1\le i \le n$. 
It is clear from the definition of $V_0$ that $\tj(i), \Phi r_j(i) \geq\Phi r_i(i)\geq V_0(i)$, $1\le i,j \le n$. 
Also from the monotone property of $T$, we have $T\Phi r_j\geq T V_0$,
hence we also have,
$\Phi r_j\geq T\Phi r_j \geq T V_0$, 
and so by taking component-wise minimum,
$V_0 \geq T V_0$.

Now let $V_{n+1} = \Gamma V_n$, $n=0,1,\dots$. Then, $V_1(i) = (\Gamma V_0)(i) = \min\{ V(i) \,:\, V\in \F_{V_0} \}
\le \min\{ V(i) \,:\, V\ge V_0, V = \Phi r, r\in \N \} \le \min\{ V(i)\,:\, V \ge V_0, V = \Phi r_j, 1\le j \le n \} = V_0(i)$,
where in the last step we used that $r_1,\dots,r_n \in \N$.
Hence, $V_1 \le V_0$. Using~\cref{gmonotone}, we then get $V_{n+1}\le V_n$ for $n\ge 0$ and thus 
$\tj\geq V_0\geq V_1\ldots\geq \tv \geq J^*$.
\end{proof}
\input{perfbnds}

>>>>

We make the following assumption on $\N$:
\begin{assumption}
\label{ass:n}
The set $\N$ satisfies the following conditions:
\begin{enumerate}[(i)]
\item \label{ass:n1}
For any $t\in \R$, $t e_1 + \N \eqdef \{ r + t e_1 \,:\, r\in \N \} = \N$, i.e., 
$\N$ is invariant to shifts by vectors of the form $t e_1$.
\item \label{ass:n2}
For $2\le k \le n$, $\sup_{r\in \N} | e_k^\top r | <+\infty$.
\item \label{ass:n3}
For each $1\le i \le n$, there exists a solution $r_i\in \R^k$ to the ALP \eqref{alp} with $c = e_i$ such that
$r_i\in \N$.
\end{assumption}

@@

Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r_{c,J}$ to be the solution to 

@@

 \underset{r\in \R^k}{\min} &\,\, c^\top \Phi r\,,\\
 \text{s.t.}& \mb \Phi r\geq  TJ\,, \qquad r\in \N\,.

@@

@@

\item Thanks to \cref{ass:n} \eqref{ass:n1} and \eqref{ass:n2}, 
for arbitrary $J$ and arbitrary \emph{nonnegative valued} $c$,
 $\Phi r_{c,J}$ is uniquely defined.
Indeed, although $r_{c,J}$ might not be unique, but $\Phi r$ where $r$ belongs to the feasible
 region underlying \eqref{lubplp}, is bounded from below, hence  $\Phi r_{c,J}$ is uniquely defined.
It also follows that $\Gamma$ is well-defined.

@@

\item Given $J\in \R^n$, define 
\begin{align*}
\F_J\eqdef\{\,\Phi r\,|\,\Phi r\geq TJ, r\in \N\,\}\,.
\end{align*}

@@

Further, since $(\Gamma J)(i) = \min\{ V(i) \,|\, V\in \F_J \}$, it also follows that $\Gamma J \le V$ holds for any $V\in \F_J$.

@@

%\begin{definition}\label{bestproj}
%\end{definition}

@@

%As mentioned earlier, the error analysis depends on two $\max$-norm contraction operators the first of which is $\Gamma$. The important result of this section is Theorem~\ref{fxpres} and it relates the fixed point $\tv$ of $\Gamma$ to $J^*$.
In what follows, we let $r^*\in \R^k$, $\bj\in \R^n$ be defined by
\begin{align*}
r^* & \eqdef \argmin_{r\in R^k}||J^*-\Phi r||_\infty\,\\
\bj & \eqdef \Gamma J^*\,.
\end{align*}
If $r^*$ is non-unique, one may choose any minimizer.
We will now show that $\Gamma$ is a contraction and that its fixed point is close to $J^*$ as long as $\bj$ is closed to $J^*$.
The following lemma shows that this latter always holds as long as $\Phi$ is well-chosen:

@@

We now show that $\Gamma$ is a contraction. 
By \cref{maxnorm}, it suffices to show that $\Gamma$ is monotone and linear along $[\one]$-rays.
We start with monotonicity.
\begin{lemma}\label{gmonotone}
For $J_1, J_2\in \R^n$ such that $J_1\geq J_2$, we have $\Gamma J_1\geq \Gamma J_2$.
\end{lemma}
\begin{proof}
Choose any $i\in \{1,\ldots,n\}$. 
Since $J_1\geq J_2$, we have $TJ_1\geq TJ_2$.
Hence, $\F_{J_1} \subset \F_{J_2}$ and thus $(\Gamma J_1)(i) \ge (\Gamma J_2)(i)$. 
Since $i$ was arbitrary, the result follows.
\end{proof}
\begin{lemma}\label{lpsol}
Let $A\in \R^{u\times v}$, $b,c\in \R^u$ and $b_0=Ax_0$ for 
some $x_0 \in \R^v$, $\N \subset \R^v$ such that $\N =x_0+ \N$. Then
\begin{align}
\min\{c^\top Ax:Ax\geq b+b_0, x\in \N\} =\min\{c^\top A y:Ay \geq b, y \in \N \}+c^\top b_0.
\end{align}
\end{lemma}
\begin{proof}
The claim follows by the change of variables $y := x-x_0$.
\end{proof}
\begin{lemma}\label{gshift}
Let $J\in \R^n$ and $t\in \R$ be a constant.
Then, $\Gamma (J+t\one) = \Gamma J + \alpha t \one$.
\end{lemma}
\begin{proof}
Consider the $i^{th}$ linear programs associated with $\Gamma J$ and $\Gamma (J+ t\one)$. 
The result follows by using Lemma~\ref{lpsol} with $A=\Phi$, $b=TJ$, $c=e_i$, $b_0=\alpha t\mathbf{1}$ 
and $x_0=\alpha t e_1$, noting that by our assumption on $\N$, $\N = \N + \alpha t e_1$.
\end{proof}
\begin{theorem}\label{gmaxcontra}
The operator $\Gamma  \colon \R^n\ra \R^n$ obeys the $\max$-norm contraction property with factor $\alpha$.
\end{theorem}
\begin{proof}
The statement follows immediately from Lemmas~\ref{gmonotone} and~\ref{gshift}, and \cref{maxnorm}.
\end{proof}
\if0
\begin{corollary}
Let $V_0$ be arbitrary and $(V_n)$ be defined recursively by
\begin{align}\label{pvi}
V_{n+1}&=\Gamma V_n, \quad n\geq 0\,.
\end{align}
Then, $V_n$ converges to the the unique fixed point $\tv$ of $\Gamma$.
\end{corollary}
\fi
Define $\tv$ denote the unique fixed point of $\Gamma$.
\todoc{Poor choice of symbols: $\tv$ is the fixed point of $\Gamma$ and $\hat{V}$ will be the fixed point of $\tilde{\Gamma}$? Hmm.. Not a high priority to change it, but still..}
\begin{lemma}\label{gfp}
The vector $\tv$ obeys $\tv\geq T\tv$.
\end{lemma}
\begin{proof}
From the definitions we have $\tv(i) = (\Gamma \tv)(i) = (\Phi r_{e_i,\tv})(i) \ge (T\tv)(i)$, $1\le i \le n$.
\if0
Consider the $i^{th}$ linear program associated with $\Gamma \tv$. We know that $\Phi r_{e_i}\geq T \tv$, for all $i=1,\ldots, n$. The result follows from noting that $\tv$ is the unique fixed point of $\Gamma $ and that $\tv(i)=(\Phi r_{e_i,V})(i)$.
\fi
\end{proof}
Recall that $\tj$ denotes the solution to the ALP in \eqref{alp}.
The fixed point of $\Gamma$ is sandwiched between $J^*$ and $\tj$:
\begin{lemma}\label{relation1}
%Assume that $\N$ is such that if $r_i$ is
%the solution to the ALP in \eqref{alp} for $c=e_i$ then $r_i\in \N$, $1\le i \le n$.
% $\tv$, the unique fixed point of the iterative scheme \eqref{pvi}, and the solution $\tj$ to the ALP in \eqref{alp}, obey 
The relation $\tj\geq\tv\geq J^*$ holds.
\end{lemma}
\begin{proof}
Since $\tv\geq T\tv$ it follows that $\tv\geq J^*$. 
Let $r_1, r_2,\ldots, r_n$ be solutions to the ALP in \eqref{alp} for $c=e_1, e_2,\ldots,e_n$ respectively
such that $r_1,\dots,r_n\in \N$. It is possible to find such solutions thanks to \cref{ass:n} \eqref{ass:n3}.
Define $V_0\in \R^n$ by $V_0(i)=\underset{j=1,\ldots, n}{\min}(\Phi r_j)(i)$, $1\le i \le n$. 
It is clear from the definition of $V_0$ that $\tj(i), \Phi r_j(i) \geq\Phi r_i(i)\geq V_0(i)$, $1\le i,j \le n$. 
Also from the monotone property of $T$, we have $T\Phi r_j\geq T V_0$,
hence we also have,
$\Phi r_j\geq T\Phi r_j \geq T V_0$, 
and so by taking component-wise minimum,
$V_0 \geq T V_0$.

Now let $V_{n+1} = \Gamma V_n$, $n=0,1,\dots$. Then, $V_1(i) = (\Gamma V_0)(i) = \min\{ V(i) \,:\, V\in \F_{V_0} \}
\le \min\{ V(i) \,:\, V\ge V_0, V = \Phi r, r\in \N \} \le \min\{ V(i)\,:\, V \ge V_0, V = \Phi r_j, 1\le j \le n \} = V_0(i)$,
where in the last step we used that $r_1,\dots,r_n \in \N$.
Hence, $V_1 \le V_0$. Using the monotone property of $\Gamma$ (i.e., \cref{gmonotone})
we then get $V_{n+1}\le V_n$ for $n\ge 0$ and thus 
$\tj\geq V_0\geq V_1\ldots\geq \tv \geq J^*$.
\end{proof}
\input{perfbnds}

### 2015-06-14 mdp.tex
LUB cleanup

%\subsection{Properties of $T$}
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$. 
Further, if $J\in \R^n$ is such that $J\geq TJ$ then $J\geq J^*$. \todoc{Should we move this second part out?}
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$ and $t \in \R$, we have
\begin{align}\label{eq:shift}
T(J+t\one)=TJ+\alpha t\one.
\end{align}
\end{lemma}
From these result, it immediately follows that $T$ is a contraction in the $L_\infty$-norm through the following lemma:
\begin{lemma}\label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{eq:shift} then 
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha$, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{proof}
The proof is given as we will need this lemma later.
Let $\eps = \norm{J_1 - J_2}_\infty$. Then $J_2 - \eps \one \le J_1 \le J_2 + \eps \one$. By the monotonicity of $T$,
$T(J_2 - \eps \one) \le T J_1 \le T(J_2 + \eps \one)$. Using~\eqref{eq:shift}, we get 
$TJ_2 - \alpha \eps \one \le T J_1 \le TJ_2 + \alpha \eps \one$, i.e., $-\alpha \eps \one \le T J_1 - T J_2 \le \alpha \eps \one$, from which the result follows.
\end{proof}
Through Banach's fixed point theorem, from \cref{maxnorm} we get the following lemma:
\begin{lemma}\label{uniquesol} 
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{lemma}
Though in these results, we make use of the Bellman operator $T$, the results also hold for $T_u$ and $H$ as well.
%Note that Lemmas~\ref{maxnorm}--\ref{shift} also hold for the Bellman operator $H$ defined for the action values in Definition~\ref{bellactval}.\par

>>>>

%\subsection{Properties of $T$}
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$. 
Further, if $J\in \R^n$ is such that $J\geq TJ$ then $J\geq J^*$. \todoc{Should we move this second part out?}
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$ and $t \in \R$, we have
\begin{align}\label{eq:shift}
T(J+t\one)=TJ+\alpha t\one.
\end{align}
\end{lemma}
Operators that satisfy \eqref{eq:shift} are said to be \emph{linear along $[\one]$-rays}.
From these result, it immediately follows that $T$ is a contraction in the $L_\infty$-norm through the following lemma:
\begin{lemma}\label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{eq:shift} then 
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha$, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{proof}
The proof is given as we will need this lemma later.
Let $\eps = \norm{J_1 - J_2}_\infty$. Then $J_2 - \eps \one \le J_1 \le J_2 + \eps \one$. By the monotonicity of $T$,
$T(J_2 - \eps \one) \le T J_1 \le T(J_2 + \eps \one)$. Using~\eqref{eq:shift}, we get 
$TJ_2 - \alpha \eps \one \le T J_1 \le TJ_2 + \alpha \eps \one$, i.e., $-\alpha \eps \one \le T J_1 - T J_2 \le \alpha \eps \one$, from which the result follows.
\end{proof}
Through Banach's fixed point theorem, from \cref{maxnorm} we get the following lemma:
\begin{lemma}\label{uniquesol} 
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{lemma}
Though in these results, we make use of the Bellman operator $T$, the results also hold for $T_u$ and $H$ as well.
%Note that Lemmas~\ref{maxnorm}--\ref{shift} also hold for the Bellman operator $H$ defined for the action values in Definition~\ref{bellactval}.\par
